# Some Non-Parametric Procedures

```{r,message=FALSE,warning=FALSE}
library(tidyverse)
library(survival)
library(readxl)
library(muhaz)
library(rms)
library(bshazard)
```


Look at the methods at the following pate to estimate S(t) using the actuarial method and the Kaplan Meier method:

http://sphweb.bumc.bu.edu/otlt/MPH-Modules/BS/BS704_Survival/BS704_Survival_print.html

## 2.1.2

The Kaplan Meier estimator of a survival curve is a step function, continuous from the right. At $t_0=0$ we have $S(t_0)=1$. Let the first survival time be $t_1$ and assume $d_1$ events where $n_1$ were at risk. This means that $d_1$ individuals had survival time equal to $t_1$. 

Then the event probability is estimated as $\hat{q}_1 = d_1/n_1$ and the survival probability as $\hat{p}_1=1-\hat{q}_1 = 1-d_1/n_1$. Then
$$
\hat{S}(t_1)=(1-\hat{q}_1)=\left(1-\frac{d_1}{n_1}\right).
$$
Assume $d_2$ events at time $t_2>t_1$ where $n_2$ are at risk.Since $S(t_2)=P[T > t_2] = P[T > t_2|T>t_1] P[T>t_1]$, it is natural to continue with $\hat{S}(t_2)= (1-\hat{q}_2)\hat{S}(t_1)$. In other words:

$$
\hat{S}(t_2)=(1-\hat{q}_2)(1-\hat{q}_1)=\left(1-\frac{d_2}{n_2}\right)\left(1-\frac{d_1}{n_1}\right).
$$

Then similarly for $t_3$, $t_4$ etc so that for any $t$
$$
\hat{S}(t) = \prod_{t_i \le t} (1-\hat{q_î}) = \prod_{t_i \le t} \left(1-\frac{d_i}{n_i}\right).
$$

Recall that $P[A | B] = P[A \cap B]/P(B)$ and note that $P[T > t_2]=P[T > t_2 \cap T > t_1]$, for $t_2>t_1$.

There is only a change in $\hat{S}(t)$ when $t$ is a time of an event. It does not change when $t$ is a censored time. However, the number at risk changes.

## 2.1.3 Nelson - Aalen

In an interval from $t_{i-1}$ to $t_i$ it is natural to estimate the hazard with the number of events in that interval divided by the person years in that interval.

$$
h(t_i) = \frac{d_i}{n_i (t_i-t_{i-1})}.
$$


The cumulative hazard can then be estimated by numerical integration:
$$
H(t) = \sum_{t_i \le t}  \frac{d_i}{n_i (t_i-t_{i-1})} (t_i-t_{i-1})
$$
so that

$$
H(t) = \sum_{t_i \le t}  \frac{d_i}{n_i}.
$$

From $\exp(a+b) = \exp(a)\exp(b)$ we get the Nelson-Aalen estimator of $S(t)$ as

$$
S(t) = \exp(-H(t)) = \prod_{j=1}^{k} \exp(-d_i/n_i)
$$
where $t_k$ is the largest time less than $t$.

## 2.2 

We can use the status method to estimate the sampling error in $\hat{S}(t)$. The **status method** is used to estimate the variance of a function of a random variable. Let $f$ be differentiable. It says that if $E(X)=\mu$ then

$$
\mathrm{Var}(f(X)) \approx \mathrm{Var(X)}\left(f^\prime(\mu)\right)^2
$$

In particular
$$
\mathrm{Var}(\log(X)) \approx \mathrm{Var(X)}(1/\mu)^2 
$$
and
$$
\mathrm{Var}(\exp(X)) \approx \mathrm{Var(X)}(\exp(\mu))^2.
$$

For example,

$$
\mathrm{var}(\log(1-\hat{q}))=\mathrm{var}(\log(\hat{p}))) \approx \mathrm{var}(\hat{p})\left(1/{p}\right)^2=\frac{\hat{p}(1-\hat{p})}{n{p}^2},
$$
which we estimate by
$$
\frac{\hat{p}(1-\hat{p})}{n\hat{p}^2}
$$


Simplifying further leads to:
$$
\mathrm{var}(\log(1-\hat{q})) \approx \frac{1-\hat{p}}{n \hat{p}} = \frac{\hat{q}}{n(1-\hat{q})} = \frac{d/n}{n(1-d/n)}=\frac{d}{n(n-d)}.
$$

Therefore 
$$
\mathrm{Var}(\log(\hat{S}(t))) = \sum_{t_i\le t} \frac{d_i}{n_i(n_i-d_i)} 
$$

Then we can write $\hat{S}(t)=\exp(\log(\hat{S}(t)))$ and use the status-method on the random variable $\log(\hat{S}(t))$ with the function $f(u)=\exp(u)$.

$$
\mathrm{Var}(\hat{S}(t)) = \mathrm{Var}(\exp(\log(\hat{S}(t)))) = \mathrm{Var}(\log(\hat{S}(t))) \left(\hat{S}(t)) \right)^2.
$$

In other words, using the results above, 
$$
\mathrm{Var}(\hat{S}(t)) =  \left(\hat{S}(t)) \right)^2 \sum_{t_i\le t} \frac{d_i}{n_i(n_i-d_i)}.
$$

Further details can be found in the [Appendix](http://onlinelibrary.wiley.com/doi/10.1002/9780470258019.app1/pdf) of the survival book by Hosmer, Lemeshow, and May.


## 2.2 Estimating the hazard function

We learn a lot about the data by studying the hazard functions.


But first, let's look at the survival. We use the data from example 1.3.


```{r}
Survival_of_multiple_myeloma_patients <- read.table("Data/Survival of multiple myeloma patients.dat", header = T)
d <- Survival_of_multiple_myeloma_patients %>% as_tibble
```


```{r}
result.km <- survfit(Surv(time, status) ~ 1,conf.type="log-log",data=d)

result.na <- survfit(Surv(time, status) ~ 1,conf.type="log-log",type="fh",data=d)
```

Kaplan Meier
```{r}
plot(result.km)
result.km
summary(result.km)
```

Nelson-Aalen
```{r}
plot(result.na)
result.na
summary(result.na)
```


In R we have the *muhaz* and *pehaz* functions to estimate the hazard.

```{r}
attach(d)
result.pe12 <- pehaz(time, status, width=12)
result.smooth <- muhaz(time, status)
detach(d)
```


The output from *pehaz*:
```{r}
result.pe12
```

The hazard in interval $i$ is estimated with $d_i / \mathrm{Follow-up}_i$ where $\mathrm{Follow-up}$ is the Total Follow-up Time  $i$.  

If we take 12 months as a fixed time interval, we get on one hand from $d_i/n_i$ 

```{r}
result.pe12$Events/(result.pe12$At.Risk*12)
```

and on the other hand the correct result.
```{r}
result.pe12$Hazard
result.pe12$Events/result.pe12$F.U.Time
```

Note that using number at risk times 12 leads to overestimate in the person-years and therefore underestimate in the hazard.

This difference gets smaller with smaller intervals. 

Graphical comparison
```{r}
attach(d)
result.pe6 <- pehaz(time, status, width=6)
detach(d)

plot(result.pe6,xlim=c(0,80),ylim=c(0,0.1))
lines(result.pe12)
points(result.pe6$Cuts[-1],result.pe6$Hazard)
lines(result.smooth$est.grid,result.smooth$haz.est,col="blue")
```

This is a comparison of the Nelson Aalen estimator (line) with the result from *pehaz*. 

Note $H(t)=-\log(S)$ is on the y-axis.

```{r}
plot(result.na$time,-log(result.na$surv))
lines(result.pe6$Cuts[-1],cumsum(result.pe6$Hazard)*6)
```

Comparison using the Kaplan-Meier results and transformations based pehaz (blue line).
```{r}
plot(result.km,xlim=c(0,90))
lines(result.pe6$Cuts[-1],exp(-cumsum(result.pe6$Hazard)*6),col="blue")
```

Results from the *rms* package
```{r}
rms1 <- npsurv(Surv(time, status) ~ 1,data=d)
```

```{r}
survplot(rms1,what=c("survival"),n.risk = T)
```

## An Application to Cat Adoptions

### Data

We'll work with the `AustinCats` data from the `rethinking` package, the companion package to the book *Statistical Rethinking* by Richard McElreath.

```{r}
data("AustinCats", package = "rethinking")
d <- AustinCats %>% 
  as_tibble %>% 
  filter(out_event %in% c("Adoption", "Censored")) %>% 
  select(days_to_event, out_event)
rm(AustinCats)
```

### Kaplan-Meier

In the *Kaplan-Meier* estimate, the survivor function is estimated and then transformed to the hazard function.

```{r, fig.asp = 0.621}
km <- d %>% 
  mutate(n = n()) %>% 
  group_by(days_to_event, n) %>% 
  summarise(events = sum(out_event == "Adoption"),
            censored = n() - events,
            lost = events + censored) %>% 
  ungroup %>% 
  mutate(total_lost = cumsum(lost),
         n = n - lag(total_lost, default = 0),
         conditional_survival = (n - events) / n,
         surv = cumprod(conditional_survival),
         cumulative_hazard = -log(surv),
         hazard = cumulative_hazard - lag(cumulative_hazard, default = 0)) %>% 
  rename(Hazard = hazard, "Cumulative Hazard" = cumulative_hazard, "Survival" = surv) %>% 
  pivot_longer(c(Hazard, "Cumulative Hazard", Survival)) %>% 
  mutate(type = "Kaplan-Meier") 

km %>% 
  ggplot(aes(days_to_event, value)) +
  geom_step() +
  facet_wrap("name", scales = "free") +
  labs(x = "Time (days)",
       title = "Kaplan-Meier estimates for cat adoptions") +
  theme(axis.title.y = element_blank())
```

### Nelson-Aalen

In the *Nelson-Aalen* estimate, the hazard function is estimated and then transformed to the survivor function.

```{r, fig.asp = 0.621}
na <- d %>% 
  mutate(n = n()) %>% 
  group_by(days_to_event, n) %>% 
  summarise(events = sum(out_event == "Adoption"),
            censored = n() - events,
            lost = events + censored) %>% 
  ungroup %>% 
  mutate(total_lost = cumsum(lost),
         n = n - lag(total_lost, default = 0),
         hazard = events / n,
         cumulative_hazard = cumsum(hazard),
         surv = exp(-cumulative_hazard)) %>% 
  rename(Hazard = hazard, "Cumulative Hazard" = cumulative_hazard, "Survival" = surv) %>% 
  pivot_longer(c(Hazard, "Cumulative Hazard", Survival)) %>% 
  mutate(type = "Nelson-Aalen")

na %>%   
  ggplot(aes(days_to_event, value)) +
  geom_step() +
  facet_wrap("name", scales = "free") +
  labs(x = "Time (days)",
       title = "Nelson-Aalen estimates for cat adoptions") +
  theme(axis.title.y = element_blank())
```

### Compare

```{r, fig.asp = 1}
km %>% 
  bind_rows(na) %>% 
  ggplot(aes(days_to_event, value)) +
  geom_step() +
  facet_grid(name ~ type, scales = "free") +
  labs(x = "Time (days)",
       title = "Comparison of Kaplan-Meier and Nelson-Aalen estimates",
       subtitle = "Based on cat adoption data from Statistical Rethinking by Richard McElreath") +
  theme(axis.title.y = element_blank())
```

